Molecular dynamics simulation of synchronization in driven particles

### 2021-01-05 manuscript/manuscript_ajp_2020.tex
fixed some language on driving force/mode locking

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient effects.
We illustrate this force in Fig.~\ref{fig:0}(a).

@@

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position % as a function of time.
for 
%We drive
a single particle across a periodic landscape. 
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$ force units 
%in simulation units for force. % $E_0/a_0$.
%The substrate is described in 
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and spatial period $\lambda = 1.825 a_0$.
%
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$.
To avoid quenching the system, 
the dc driving force is initially zero   
and 
slowly increased %$F^{DC}$
at a rate of $\Delta F^{DC} = 0.01$ every $\Delta t = 4000$ integer timesteps 
to the constant value $F^{DC}=0.1$,
as shown in the blue curve in Fig.~\ref{fig:0}(a)
where $F^{DC}$ increases for the first $40$ time units.
The combined ac and dc drive $F^D(t)$ is shown in orange,
where 
the magnitude of the ac applied force is initially zero
since $\sin{(0)}=0$.
The vertical dashed lines in panels (a) and (b) coincide
with the condition $\sin{(\omega t)} = 1$,
where the applied force is maximum once $F^{DC}$ has reached its constant value.

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by the spatial period of the potential %washboard as
$y/\lambda$.
The initial particle position is $y=0$,
where the landscape force is minimum $F^l = 0$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum 
$\sin{(\pi y / \lambda)} = 1 $. 
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
The crossing of the particle position $y$
with the horizontal and vertical lines
indicates the synchronization
of the driving force with the lattice position of the particle,
where driving force is maximum when the particle
is at the landscape minimum.

When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape troughs,
and 
the particle hops between minima in the energy landscape.
{\it UPDATE:} 
In Fig.~\ref{fig:0} we demonstrate this relationship,
where over half of a driving period ($T = 50$) %time units
while $F_{D}(t) > 0 $ 
the particle moves
in the positive y-direction
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.

>>>>

 Particles are subject an external time-dependent driving force
$\vec{F}_{D}(t)$
applied parallel to the y-direction.
We model this force as
\begin{equation}
  \vec{F}^{D}(t) = [F^{DC} + F^{AC} \sin(\omega t)] \hat{y},
    \label{eq:drive}
\end{equation}
with modifiable parameters including
a constant component $F^{DC}$,
and a time dependent component with amplitude $F^{AC}$
and frequency $\omega = 2 \pi f$.
%in the positive $y-$direction.
%but it may 
%be zero or negative depending the the parameters and time.
When $F^{DC}$ is non-zero,  
%In several simulations
we slowly increase 
$F^{DC}$ from zero
to avoid transient behavior
by ensuring the particle is subject to a smoothly varying force.
We illustrate the increase
of $F^{DC}$ in Fig.~\ref{fig:0}(a).

@@

In Fig.~\ref{fig:0} we
show the relationship between the applied force and particle
position % as a function of time.
for 
%We drive
a single particle driven across a periodic landscape. 
%along the y-direction
%({\it redundant, just refer back to eq. in Model?})
%and $V_{0y}=0.2$.
%This is illustrated in Fig.~\ref{fig:1}(a) where
%the red (blue) regions show local maxima (minima).
%
The numerical implementation of the landscape 
is calculated with Eq.~\ref{eq:ysubstrate} as 
\begin{equation}
  \label{eq:force}
  F^l_y(y) = -A_{p} \sin{(N_p \pi y / L)} 
\end{equation}
where the force is scaled with parameter $A_{p} = 0.1$ force units 
%in simulation units for force. % $E_0/a_0$.
%The substrate is described in 
%\begin{equation}
%  \label{eq:landscape_potential}
%  U_l(y) = U_{0y} \cos{(N_p \pi y / L)} 
%\end{equation}
with $N_p=20$ %troughs in the landscape
and spatial period $\lambda = 1.825 a_0$.
%
In Fig.~\ref{fig:0}(a)
we plot the applied driving force defined in Eq.~\ref{eq:drive}
as a function of time.
The time independent constants are 
$F^{AC}=0.2$ and $f=0.01$ cycles per time unit.
The temporal period of the driving force is
$T = 1/f = 100$.
To avoid transient oscillations at early times, 
the dc driving force is initially zero   
and 
slowly increased %$F^{DC}$.
at a rate of $\Delta F^{DC} = 0.01$ every $\Delta t = 4000$ integer timesteps 
to the constant value $F^{DC}=0.1$.
In Fig.~\ref{fig:0}(a) we plot $F^{DC}$
in blue to show the increase in 
$F^{DC}$ for the first $40$ time units.
In orange we show 
the combined ac and dc drive $F^D(t)$,
where 
the magnitude of the ac applied force is initially zero.
%since $\sin{(0)}=0$.
The vertical dashed lines in panels (a) and (b) coincide
with the condition $\sin{(\omega t)} = 1$,
where the applied force is maximum once $F^{DC}$ has become constant.

@@

In Fig.~\ref{fig:0}(b) 
we show the $y-$position of the particle
as a function of time,
where we 
normalize $y$ by the spatial period of the potential %washboard as
$\lambda$.
The initial particle position is $y=0$,
where the landscape force is minimum $F^l = 0$.
The horizontal dashed lines
coincide with the condition the substrate force is
minimum 
$\sin{(\pi y / \lambda)} = 1 $, so $F^l = -A_p$.
%or $\pi y / \lambda = m \pi/2$, where $m = 1, 5, 7$, {\it etc.}
The crossing of the particle position $y$
with the horizontal and vertical lines
indicates the synchronization
of the driving force with the lattice position of the particle,
where driving force is maximum positive when the
force on the particle from the landscape is maximum negative.
It can be shown this synchronization
occurs for any initial position of $y$.

The slope of $y/\lambda$
is proportional to the net force on the particle (See Eq.~\ref{eq:motion}).
This can be seen in 
Fig.~\ref{fig:0} %we demonstrate this relationship,
where local extrema appear in $y/\lambda$
when $F_{D}(t) \approx 0 $ and 
When $F^D(t) > A_p$, a particle can 
overcome the barrier height of the landscape maxima,
and 
the particle hops between minima in the energy landscape.

the particle moves
in the positive y-direction
through three
substrate troughs, reaching a maximum of $y/\lambda = 3$.
When $F_{D}(t) < 0 $
the particle moves in the negative y-direction
reaching a minimum position
of $y/\lambda = 1$
over the total period $T = 100$.
The average velocity $\bar{v}_y$
is the average displacement $\Delta y = y(t_0+T) - y(t_0)$
over the period of the driving force.
In Fig. ~\ref{fig:0}
the average displacement during one time period is a single wavelength
of the substrate $y(t_0+T) - y(t_0) = \lambda$.
Thus the average velocity 
is $\bar{v}_y = \lambda f$. %, where $T = 1/f$.
